---
title: "ODD Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Serhan SÜER


## From Raw to Civilized Data

First we find the data on Otomotiv Distibütörleri Derneği [website](http://www.odd.org.tr/web_2837_1/neuralnetwork.aspx?type=36). We are interested in January 2018 sales. We download the data change the name to odd_retail_sales_2018_01.xlsx.


## Download Raw Data

Our raw excel file is in our repository. We can automatically download that file and put it in a temporary file. Then we can read that excel document into R and remove the temp file.


```{r}
# Create a temporary file
tmp<-tempfile(fileext=".xlsx")
# Download file from repository to the temp file
download.file("https://github.com/MEF-BDA503/pj18-SerhanSuer/blob/master/odd_retail_sales_2018_01.xlsx?raw=true",destfile=tmp)
# Read that excel file using readxl package's read_excel function. You might need to adjust the parameters (skip, col_names) according to your raw file's format.
raw_data<-readxl::read_excel(tmp,skip=7,col_names=FALSE)
# Remove the temp file
file.remove(tmp)
```

```{r include = FALSE}
library(tidyverse)
```


```{r}
# Remove the last two rows because they are irrelevant (total and empty rows)
raw_data <- raw_data %>% slice(-c(43,44))
# Let's see our raw data
head(raw_data)
```


## Make Data Civilized

In order to make the data standardized and workable we need to define column names and remove NA values for this example.

```{r}
# Use the same column names in your data.
colnames(raw_data) <- c("brand_name","auto_dom","auto_imp","auto_total","comm_dom","comm_imp","comm_total","total_dom","total_imp","total_total")
# Now we replace NA values with 0 and label the time period with year and month, so when we merge the data we won't be confused.
car_data_jan_18 <- raw_data %>% 
    mutate_if(is.numeric,funs(ifelse(is.na(.),0,.))) %>% 
    mutate(year=2018,month=1)

print(car_data_jan_18,width=Inf)
```


## Finish With Some Analysis
I wanted to make four analysis related to this dataset.


1. Firstly, I wanted to see a list of total sales of brands with both automobile and commercial vehicle sales ordered only in domain sector in decreasing total sales.
```{r}
car_data_jan_18 %>% 
    filter(total_imp == 0 & total_total != 0) %>%
    select(brand_name,total_total) %>%
    arrange(desc(total_total))
```

2. Then I wanted to see a list of total sales of brands with both automobile and commercial vehicle sales ordered only in import sector in decreasing total sales.
```{r}
car_data_jan_18 %>% 
    filter(total_dom == 0 & total_total != 0) %>%
    select(brand_name,total_total) %>%
    arrange(desc(total_total))
```

3. Then I wanted to see a list of total sales of brands with only automobile vehicle sales ordered in decreasing total sales.
```{r}
car_data_jan_18 %>% 
    filter(comm_total == 0 & total_total != 0) %>%
    select(brand_name,total_total) %>%
    arrange(desc(total_total))
```

4. Then I wanted to see a list of total sales of brands with only commercial vehicle sales ordered in decreasing total sales.
```{r}
car_data_jan_18 %>% 
    filter(auto_total == 0 & total_total != 0) %>%
    select(brand_name,total_total) %>%
    arrange(desc(total_total))
```